---
title: "QMM_Integer_Assignment"
author: "Dilip Kumar"
date: "2023-11-20"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(lpSolve)

# Creating a matrix for days and required workers
DAY_AND_WORKERS<- matrix(c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday",18,27,22,26,25,21,19),ncol = 2, byrow = F) 
colnames(DAY_AND_WORKERS) = c("Day","Workers_Required")

# Converting to a table for display
as.table(DAY_AND_WORKERS)
```





Package handlers at AP are guaranteed a five-day work week with two consecutive days off. The
base wage for the handlers is $750 per week. Workers working on Saturday or Sunday receive an
additional $25 per day. The possible shifts and salaries for package handlers are:

```{r}

# Making a table for shifts, days off, and wages
Shift_DaysOff_Wage <- matrix(c(1,2,3,4,5,6,7,"Sunday and Monday","Monday and Tuesday","Tuesday and Wednesday","Wednesday and Thursday","Thursday and Friday","Friday and Saturday","Saturday and Sunday","$775","$800","$800","$800","$800","$775","$750"), ncol = 3, byrow = F)

# Adding column names
colnames(Shift_DaysOff_Wage) <- c("Shift", "Days_Off", "Wage")

# Converting to a table for display
as.table(Shift_DaysOff_Wage)


```


Question:
The manager wants to keep the total wage expenses as low as possible while ensuring that there
are sufficient number of workers available each day. Formulate and solve the problem. What was
the total cost? How many workers are available each day?

Solution:
##Formulate and solve LP problem


#Objective Function:
We aim to minimize the total cost, which is represented by the objective function:

Z = 775X_1 + 800X_2 + 800X_3 + 800X_4 + 800X_5 + 775X_6 + 750X_7

Where:
(X1, X2, X3, X4, X5, X6, X7) are the number of workers under specific shift schedules with corresponding salaries.

#Constraints:
We have several constraints to ensure there are sufficient workers available each day:

1. Availability on Sunday:    0X_1 + 1X_2 + 1X_3 + 1X_4 + 1X_5 + 1X_6 + 0X_7 

2. Availability on Monday:    0X_1 + 0X_2 + 1X_3 + 1X_4 + 1X_5 + 1X_6 + 1X_7 

3. Availability on Tuesday:   1X_1 + 0X_2 + 0X_3 + 1X_4 + 1X_5 + 1X_6 + 1X_7 

4. Availability on Wednesday: 1X_1 + 1X_2 + 0X_3 + 0X_4 + 1X_5 + 1X_6 + 1X_7 

5. Availability on Thursday:  1X_1 + 1X_2 + 1X_3 + 0X_4 + 0X_5 + 1X_6 + 1X_7 

6. Availability on Friday:    1X_1 + 1X_2 + 1X_3 + 1X_4 + 0X_5 + 0X_6 + 1X_7 

7. Availability on Saturday:  1X_1 + 1X_2 + 1X_3 + 1X_4 + 1X_5 + 0X_6 + 0X_7

#Non-Negative Constraints:
To ensure a realistic solution, we have non-negative constraints for each variable: [ X_{1:7} ]

These constraints collectively represent the optimization problem you are solving—minimizing the total cost while meeting the daily worker availability requirements.




Let's solve the problem using "R", and by using above mentioned library "lpsolve".
Since, this problem needs to be solved by integer programming, we need to use the function int.vec, which says that all the variables should be of integers and not fractions.

```{r}

# Define the objective function coefficients
obj.f <- c(775, 800, 800, 800, 800, 775, 750)

# Defining the constraints matrix
const <- matrix(c(0, 1, 1, 1, 1, 1, 0,
                 0, 0, 1, 1, 1, 1, 1,
                 1, 0, 0, 1, 1, 1, 1,
                 1, 1, 0, 0, 1, 1, 1,
                 1, 1, 1, 0, 0, 1, 1,
                 1, 1, 1, 1, 0, 0, 1,
                 1, 1, 1, 1, 1, 0, 0), nrow = 7, byrow = TRUE)

const

```





```{r}
row.signs <- rep(">=", 7)

# Defining the right-hand side of constraints
row.rhs <- c(18, 27, 22, 26, 25, 21, 19)
integer.program_cost<-lp("min",obj.f,const,row.signs,row.rhs,int.vec=1:9)
integer.program_cost
```
```{r}
# Displaying the number of workers for each shift

integer.program_cost$solution#This show how many workers are corresponding to each shift on respective days
```

The weekly salary total is $25,675.



These represent the total employees, along with their salaries and shift schedules, as shown below:
```{r}
#By using the variables from the lp model 
Shift_table = matrix(c(0,4,5,0,8,1,0,0,0,5,0,8,1,13,2,0,0,0,8,1,13,2,4,0,0,8,1,13,2,4,5,0,0,1,13,2,3,4,0,0,0,13,2,4,5,0,8,0,0),ncol=7,byrow=TRUE)

colnames(Shift_table) = c("Shift1", "Shift2", "Shift3", "Shift4", "Shift5", "Shift6", "Shift7")

row.names(Shift_table) = c('Sunday', 'Monday', 'Tuesday','Wednesday','Thursday','Friday','Saturday')

print(Shift_table)
```

The Shift_table below illustrates the number of available employees for each day, strategically arranged to minimize overall wage expenses.

```{r}
# Showing the sum of workers present for each day.
rowSums(Shift_table)
```
